\documentclass[12pt, a4paper]{article}

\usepackage[utf8]{inputenc}
% Limit the page margin to only 1 inch.
\usepackage[margin=1in]{geometry}

%Imports biblatex package
\usepackage[
backend=biber,
style=alphabetic
]{biblatex}
\addbibresource{../../algs4e.bib}

\usepackage{amsmath}
\usepackage{amssymb}

% Provides useful environments, such as:
% - \begin{proof} ...\end{proof}
\usepackage{amsthm}

\usepackage[inline]{enumitem} % allows using letters in enumerate list environment

% Enable importing external graphic files and provides useful commannds, like \graphicspath{}
\usepackage{graphicx}
% Images are located in a directory called images in the current directory.
\graphicspath{{./images/}}

% Code Listings. Source:
% https://stackoverflow.com/questions/3175105/inserting-code-in-this-latex-document-with-indentation
\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
	language=Java,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numbers=none,
	numberstyle=\tiny\color{gray},
	keywordstyle=\color{blue},
	commentstyle=\color{dkgreen},
	stringstyle=\color{mauve},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3
}

\newcommand{\prob}{\text{P}}
%\newcommand{\complement}{\mathsf{c}}

% Define an environment called "ex" (for Exercise) so that I can do: \begin{ex}{1.5}...\end{ex}
\newenvironment{ex}[2][Exercise]
{\par\medskip\noindent \textbf{#1 #2.}}
{\medskip}

% Define a solution environment, similar to ex (exercise) environment.
\newenvironment{sol}[1][Solution]
{\par\medskip\noindent \textbf{#1.} }
{\medskip}

\begin{document}
	\noindent Sergio E. Garcia Tapia \hfill
	
	\noindent \emph{Algorithms} by Sedgewick and Wayne (4th edition) \cite{sedgewick_wayne}\hfill
	
	\noindent September 14th, 2024\hfill 
	\section*{1.2: Data Abstraction}
	\begin{ex}{1}
		Write a \texttt{Point2D} client that takes an integer value $n$ from the command line,
		generates $n$ random points in the unit square, and computes the distance separating the
		\emph{closest pair} of points.
	\end{ex}
	\begin{sol}
		\
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex01.ClosestPointPair} class.
		The \texttt{Point2D} objects can simply by stored in an array of size $n$.
		We can use \texttt{StdRandom.uniformDouble()} to generate random $x$ and $y$ coordinates
		for each point, and then leverage the \texttt{distanceTo()} method available in the
		\texttt{Point2D} API. I employed a nested \texttt{for} loop to compute the closest pair.
	\end{sol}
	\begin{ex}{2}
		Write an \texttt{Interval1D} client that takes an \texttt{int} value $n$ as command-line
		argument, reads $n$ intervals (each defined by a pair of \texttt{double} values)
		from standard input, and prints all pairs that intersect.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex02.IntervalIntersection} class.
		It's much the same as in Exercise 1, but instead of using \texttt{StdRandom.uniformDouble()}
		to generate the coordinates, I use \texttt{StdIn.readDouble()} to read coordinate from
		standard input. Also, instead of the \texttt{distanceTo()} method from the
		\texttt{Point2D} API, I leveraged the \texttt{intersects()} method from the
		\texttt{Interval1D} API.
	\end{sol}
	\begin{ex}{3}
		Write an \texttt{Interval2D} client that takes command-line arguments \texttt{n},
		\texttt{min}, and \texttt{max} and generates \texttt{n} random 2D intervals whose
		width and height are uniformly distributed between \texttt{min} and \texttt{max}
		in the unit square. Draw them on \texttt{StdDraw} and print the number of pairs
		of intervals that intersect and the number of intervals that are contained in one
		another.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex03.IntersectingRectangles} class.
		
		One important consideration is that since the widths and heights are generated uniformly
		between \texttt{min} and \texttt{max}, we must ensure the bottom left corner of each
		point isn't so large that it would exceed the  dimensions of the unit square.
		That is, given \texttt{width} and \texttt{height}, each of the $x$ and $y$
		coordinates of the bottom left vertex of all rectangles must not exceed
		\texttt{1 - width} and \texttt{1 - height}, respectively.
		
		Another important consideration is that, to check if rectangle $A$ contains rectangle $B$,
		we must check that the bottom-left and top-right vertices of rectangle $B$ are contained in
		rectangle $A$.Since the \texttt{Interval2D} API does not expose methods for obtaining these
		quantities, it's necessary to save them while doing the computations to necessary to create
		the rectangles.
	\end{sol}
	\begin{ex}{4}
		What does the following code fragment print?
		\begin{lstlisting}
		String string1 = "hello";
		String string2 = string1;
		string1 = "word";
		StdOut.println(string1);
		StdOut.println(string2);
		\end{lstlisting}
	\end{ex}
	\begin{sol}
		When \texttt{string1} is assigned to\texttt{string2}, the \texttt{string2} variable
		receives a copy of the reference to the current value of \texttt{string1}.
		When \texttt{string1} is assigned the \texttt{String} with value \texttt{"word"},
		the reference in \texttt{string2} is unchanged. Thus the output is:
		\begin{lstlisting}[language={}]
world
hello
		\end{lstlisting}
	\end{sol}
	\begin{ex}{5}
		What does the following code fragment print?
		\begin{lstlisting}
		String s = "Hello World";
		s.toUpperCase();
		s.substring(6, 11);
		StdOut.println(s);
		\end{lstlisting}
	\end{ex}
	\begin{sol}
		\texttt{String} objects are immutable, so the calls to the  \texttt{toUpperCase()}
		and the \texttt{substring()} methods do not change the object that \texttt{s}
		references; they return new \texttt{String} objects. In this case, those objects
		are not stored, so they are immediately available for garbage collection. Thus the
		output is simply:
		\begin{lstlisting}[language={}]
Hello World
		\end{lstlisting}
	\end{sol}
	\begin{ex}{6}
		A string \texttt{s} is a circular shift (or \emph{circular rotation}) of a string
		\texttt{t} if it matches when the characters are circularly shifted by any number
		of positions; e.g., \texttt{ACTGACG} is a circular shift of \texttt{TGACGAC}, and
		viceversa. Detecting this condition is important in the study of genomic sequences.
		Write a program that checks whether two given strings \texttt{s} and \texttt{t}
		are circular shifts of one another. \emph{Hint}: The solution is a one liner with
		\texttt{indexOf()}, \texttt{length()}, and string concatenation.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex06.CircularShift} class.
		A prerequisite for \texttt{s} and \texttt{t} to be circular shifts of one another
		is that they have the same length. After establishing this, we can detect the
		condition by concatenating \texttt{s} with itself to create a new string,
		and then check whether \texttt{t} is a substring of this new string.
	\end{sol}
	\begin{ex}{7}
		What does the following recursive function return?
		\begin{lstlisting}
	public static String mystery(String s)
	{
		int n = s.length();
		if (n <= 1) return s;
		String a = s.substring(0, n/2);
		String b = s.substring(n/2, n);
		return mystery(b) + mystery(a);
	}
		\end{lstlisting}
	\end{ex}
	\begin{sol}
		It reverses the string \texttt{s}.
	\end{sol}
	\begin{ex}{8}
		Suppose \texttt{a[]} and \texttt{b[]} are each integer arrays consisting of millions
		of integers. What does the following code do? Is it reasonably efficient?
		\begin{lstlisting}
		int[] t  = a; a = b; b = t;
		\end{lstlisting}
	\end{ex}
	\begin{sol}
		It swaps arrays \texttt{a} and \texttt{b}. It's very efficient because the array
		values are not copied.
	\end{sol}
	\begin{ex}{9}
		Instrument \texttt{BinarySearch} (page 47) to use a \texttt{Counter} to count the total
		number of keys examined during all searches and then print the total after all searches
		are complete. \emph{Hint}: Create a \texttt{Counter} in \texttt{main()} and pass it as
		an argument to \texttt{indexOf()}.
	\end{ex}
	\begin{sol}
		See the See the \texttt{com.segarciat.algs4.ch1.sec2.ex09.BinarySearchCounter} class.
	\end{sol}
	\begin{ex}{10}
		Develop a class \texttt{VisualCounter} that allows both increment and decrement
		operations. Take two arguments \texttt{n} and \texttt{max} in the constructor,
		where \texttt{n} specifies the maximum number of operations and \texttt{max} specifies
		the maximum absolute value of the counter. As a side effect, create a plot showing the
		value of the counter each time its tally changes.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex10.VisualCounter} class.
	\end{sol}
	\begin{ex}{11}
		Develop an implementation \texttt{SmartDate} of our \texttt{Date} API that raises
		an exception if the date is not legal.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex11.SmartDate} class.
	\end{sol}
	\begin{ex}{12}
		Add a method \texttt{dayOfTheWeek()} to \texttt{SmartDate} that returns a \texttt{String}
		value \texttt{Monday}, \texttt{Tuesday}, \texttt{Wednesday}, \texttt{Thursday},
		\texttt{Friday}, \texttt{Saturday}, or \texttt{Sunday}, giving the appropriate
		day of the week for the date. You may assume that the date is in the 21st  century.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex12.SmartDate} class.
	\end{sol}
	\begin{ex}{13}
		Using our implementation of \texttt{Date} as a model (page 91), develop an implementation
		of \texttt{Transaction}.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex13.Transaction} class.
	\end{sol}
	\begin{ex}{13}
		Using our implementation of \texttt{equals()} in \texttt{Date} as a model (page 103),
		develop an implementation of \texttt{equals()} for \texttt{Transaction}.
	\end{ex}
	\begin{sol}
		See the \texttt{com.segarciat.algs4.ch1.sec2.ex14.Transaction} class.
	\end{sol}
	\pagebreak
	\printbibliography
\end{document}